Well-quasi-orderings on word languages \\ main dd2237175323943c3893f9ae93107aba9f6049a5 \\ 2025-09-08 16:09:49 +0200

### 2025-09-29 wqo-on-words.lncs.tex
fix: institutions

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main 0121a5929682a015f9ca5ed9399070224366a74c \\
       2025-08-12 15:10:21 +0200
}
\titlerunning{WQO on Words}

\author{
        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
\authorrunning{N. Lhote and A. Lopez and L. Schütze}
\institute{
University of Warsaw
 \and
Aix-Marseille University
 \and
Max Planck Institute for Software Systems
}

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered. This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science. However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering. In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems. More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings. We show that a simple characterization is possible for the prefix and suffix orderings, and that under extra regularity assumptions, this also extends to the infix ordering. We furthermore provide decision procedures for a large class of languages, that contains regular and context-free languages.
\end{abstract}

>>>>

\begin{document}
%
\title{Well-quasi-orderings on word languages \\
       main dd2237175323943c3893f9ae93107aba9f6049a5 \\
       2025-09-08 16:09:49 +0200
}
\titlerunning{WQO on Words}

\author{
        Nathan Lhote\inst{2}\orcidID{0000-0003-3303-5368} \and
        Aliaume Lopez\inst{1}\orcidID{0000-0002-4205-327X} \and
        Lia Schütze\inst{3}\orcidID{0000-0003-4002-5491}}
\authorrunning{N. Lhote and A. Lopez and L. Schütze}
\institute{
INP Bordeaux, LaBRI, CNRS
 \and
Aix-Marseille University
 \and
Max Planck Institute for Software Systems
}

@@

%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    The set of finite words over a well-quasi-ordered set is itself well-quasi-ordered. This seminal result by Higman is a cornerstone of the theory of well-quasi-orderings and has found numerous applications in computer science. However, this result is based on a specific choice of ordering on words, the (scattered) subword ordering. In this paper, we describe to what extent other natural orderings (prefix, suffix, and infix) on words can be used to derive Higman-like theorems. More specifically, we are interested in characterizing \emph{languages} of words that are well-quasi-ordered under these orderings, and explore their properties and connections with other language theoretic notions. We furthermore give decition procedures when the languages are given by various comupational models such as automata, context-free grammars, and automatic structures.
\end{abstract}
